Hazelnut: A Bidirectionally Typed \\ Structure Editor Calculus

### 2016-06-28 hazelnut-popl17.tex
adding some todos

Functions are not themselves annotated with types, so they can only appear in analytic position:
\begin{equation}\label{rule:syn-lam}
\inferrule{
  \hana{\hGamma, x : \htau_1}{\hexp}{\htau_2}
}{
  \hana{\hGamma}{\hlam{x}{\hexp}}{\tarr{\htau_1}{\htau_2}}
}
\end{equation}
(It would be straightforward to also add a ``half-annotated'' lambda form, $\lambda x{:}\tau.e$, but for simplicity, we leave it out of our calculus \cite{DBLP:conf/tldi/ChlipalaPH05}.)

@@

The hole type behaves much like the type $?$ in prior work by Siek and Taha on gradual types for functional languages \cite{Siek06a}. Their system (which was not bidirectionally typed nor an editor model) also needed to define two rules for function application. In general, when a premise requires that a synthesized type be of a particular form, we need a special case where the synthesized hole type is treated instead as if it were the ``holey-est'' type of that form.\footnote{Alternatively, we might add a rule that allows expressions that synthesize hole type to then non-deterministically synthesize any other type, but maintaining determinism is useful in practice, so we avoid this approach.}

@@

As a consequence of applying Theorem \ref{thrm:actsafe} inductively at
every action in such a list, we couldn't hope to construct an ill-typed
term. Therefore the statement here matches the mutually recursive structure
of the bidirectional typing judgements.

@@

\paragraph{Finishability}\todo{do we care?}

As with deletion, since the meaning of the finish action does not depend on
the structure of the term being finished other than requiring that it must
be one of the two hole forms, there is no checksum theorem to state for
this action.

>>>>

Functions are not themselves annotated with types, so they can only appear in analytic position:
\begin{equation}\label{rule:syn-lam}
\inferrule{
  \hana{\hGamma, x : \htau_1}{\hexp}{\htau_2}
}{
  \hana{\hGamma}{\hlam{x}{\hexp}}{\tarr{\htau_1}{\htau_2}}
}
\end{equation}
It would be straightforward to also add a ``half-annotated'' lambda form, $\lambda x{:}\tau.e$, but for simplicity, we leave it out of our calculus \cite{DBLP:conf/tldi/ChlipalaPH05}.

@@

\todo{redundant text w/ citation}
The hole type behaves much like the type $?$ in prior work by Siek and Taha on gradual types for functional languages \cite{Siek06a}. Their system (which was not bidirectionally typed nor an editor model) also needed to define two rules for function application. In general, when a premise requires that a synthesized type be of a particular form, we need a special case where the synthesized hole type is treated instead as if it were the ``holey-est'' type of that form.\footnote{Alternatively, we might add a rule that allows expressions that synthesize hole type to then non-deterministically synthesize any other type, but maintaining determinism is useful in practice, so we avoid this approach.}

@@

As a consequence of applying Theorem \ref{thrm:actsafe} inductively at
every action in such a list, it is possible consider moving to an ill-typed
term. Therefore the statement here matches the mutually recursive structure
of the bidirectional typing judgements.

@@

\paragraph{Finishability}\todo{do we care?}
As with deletion, since the meaning of the finish action does not depend on
the structure of the term being finished other than requiring that it must
be one of the two hole forms, there is no checksum theorem to state for
this action.
